Use of the package \texttt{fitdistrplus} to specify a distribution from non-censored or censored data

### 2010-04-23 inst/doc/intro2fitdistrplus.Rnw
Some minor and major modifications on fitdistrplus (see NEWS)

In order to take into account the uncertainty of the estimated values of kurtosis and skewness, 
the data set may be boostrapped by fixing the argument \texttt{boot} to an integer above 10 in \texttt{descdist}. 
\texttt{boot} values of 
skewness and kurtosis corresponding to the boot nonparametric bootstrap samples are then computed and reported in blue color 
on the skewness-kurtosis plot.

@@

One or more parametric distributions may then be fitted to the data set, one at a time, using the
fonction \texttt{fitdist}. This function uses the maximum likelihood
method if the argument \texttt{method="mle"} (or if it is omitted) 
or the matching moments estimation if the argument \texttt{method="mme"}.
When fitting continuous distributions, 
Kolmogorov-Smirnov and Anderson-Darling statistics are computed and corresponding tests are performed when possible. 
Even if less appropriate for continuous distributions, the Chi-squared statistic is
also computed when possible. For this calculation, cells are defined by the argument \texttt{chisqbreaks}
or automatically defined from the data set and from the argument \texttt{meancount} (the approximate mean count per cell)
which is  fixed to $(4n)^{2/5}$ if omitted (with n the length of the data set).  
For more details, see the help of the function \texttt{fitdist}.
Four goodness of fit plots are also provided.

Below is the result of a fit of a gamma distribution by maximum likelihood.
<<fitdistx1g,fig=TRUE>>=
f1g<-fitdist(x1,"gamma")
plot(f1g)
summary(f1g)
@

Below is the result of another fit of the same distribution by matching moments.
<<>>=
f1gbis<-fitdist(x1,"gamma",method="mme")
summary(f1gbis)
@

As can be seen in this returned summary, the automatic definition of the cells required to calculate the Chi-squared
statistic does not give theoretical counts large enough to validate the use of the test in this example.
 It is often the case for small
data sets. The observed and theoretical counts may be printed as below :
<<>>=
f1g$chisqtable
@

Below is the fit of a lognormal distribution.
<<fitdistx1ln,fig=TRUE>>=
f1l<-fitdist(x1,"lnorm")
plot(f1l)
summary(f1l)
@

Below is the fit of a normal distribution.
<<fitdistx1n,fig=TRUE>>=
f1n<-fitdist(x1,"norm")
plot(f1n)
summary(f1n)
@

Below is the fit of a Weibull distribution.
<<fitdistx1w,fig=TRUE>>=
f1w<-fitdist(x1,"weibull")
plot(f1w)
summary(f1w)
@

The values of the Anderson-Darling statistic (or another result of the fit: see the help of \texttt{fitdist} for details)
 for the different fittings may be extracted and compared to help the selection of a distribution :

<<>>=
anderson<-list(lnorm=f1l$ad,gamma=f1g$ad,norm=f1n$ad,weibull=f1w$ad)
anderson
@

@@

And a Gumbel distribution may be fitted to data with these values for initial parameter values.
<<fitdistx1gu,fig=TRUE>>=
fgu<-fitdist(x1,"gumbel",start=list(a=10,b=5))
plot(fgu)
summary(fgu)
@

@@

Below is an example of the use of this function with the previous fit of the gamma distribution. 
<<bootdistf1g,fig=TRUE>>=
b1g<-bootdist(f1g)
plot(b1g)
summary(b1g)
@

@@

<<descdistx2,fig=TRUE>>=
descdist(x2,discrete=T)
@

@@

Below is the result of the fit of a Poisson distribution with the bootstrap simulations.
<<fitdistx2p,fig=TRUE>>=
f2p<-fitdist(x2,"pois")
plot(f2p)
summary(f2p)
b2p<-bootdist(f2p)
summary(b2p)
@

Below is the result of the fit of a negative binomial distribution with the boostrap simulations.
<<fitdistx2nb,fig=TRUE>>=
f2n<-fitdist(x2,"nbinom")
plot(f2n)
summary(f2n)
b2n<-bootdist(f2n)
summary(b2n)
@

>>>>

Skewness and kurtosis are known not to be robust. In order to try to
take into account the uncertainty on the estimated values of kurtosis and skewness, 
the data set may be boostrapped by fixing the argument \texttt{boot} to an integer above 10 in \texttt{descdist}. 
\texttt{boot} values of 
skewness and kurtosis corresponding to the boot nonparametric bootstrap samples are then computed and reported
on the skewness-kurtosis plot.

@@

One or more parametric distributions may then be fitted to the data set, one at a time, using the
fonction \texttt{fitdist}. This function uses the maximum likelihood
method if the argument \texttt{method="mle"} (or if it is omitted) 
or the matching moments estimation if the argument \texttt{method="mme"}.
When fitting continuous distributions, 
Kolmogorov-Smirnov and Anderson-Darling statistics may be computed using the function \texttt{gofstat}. 
Four goodness of fit plots are also provided.

Below is the result of the fit of a logistic distribution by maximum likelihood to the previous dataset.
<<fitdistx1l,fig=TRUE>>=
f1l<-fitdist(x1,"logis")
summary(f1l)
gofstat(f1l)
plot(f1l)
@

In that case the Anderson-Darling test may be used to test the adequation of the distribution to data. 
<<>>=
gofstat(f1l,print.test=TRUE)
@

Generally goodness-of-fit tests may be used carefully.
In somme cases, especially on very big datasets, even if the null hypothesis is rejected a fitted distribution may be 
chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
plots do not show strong differences between empirical and theoretical distributions.

In that case, the fit seems correct, but it is easy to compare goodness-of-fit statistics
for other distributions fitted on the same dataset in order to check that other distributions could not give
a better fit. Below are computed the goodness-of-fit
statistics for logistic, lognormal, gamma, normal and weibull distributions. 
<<>>=
gofstat(fitdist(x1,"logis"))
gofstat(fitdist(x1,"lnorm"))
gofstat(fitdist(x1,"gamma"))
gofstat(fitdist(x1,"norm"))
gofstat(fitdist(x1,"weibull"))
@

Regarding the Kolmogorov-Smirnov statistic, the fit of a normal distribution seems better while regarding the 
Anderson-Darling statistic the fit of a logistic distribution seems better.
It is recommended to base the comparison of fits on the Anderson-Darling statistic when the modelling of the tails of 
a distribution is important, as it is often the case in risk assessment.

In order to choose between both distributions, the goodness-of-fit plot and the summary
of the fit of a normal chould be compared to the 
ones already obtained for the logistic distribution, but in that case the differences seem very small.
<<fitdistx1n,fig=TRUE>>=
f1n <- fitdist(x1,"norm")
plot(f1n)
summary(f1n)
plot(f1n)
@

@@

And a Gumbel distribution may be fitted to data with these values for initial parameter values.
<<fitdistx1gu,fig=TRUE>>=
fgu<-fitdist(x1,"gumbel",start=list(a=10,b=5))
plot(fgu)
summary(fgu)
gofstat(fgu)
@

@@

Below is an example of the use of this function with the previous fit of the logistic distribution. 
<<bootdistf1l,fig=TRUE>>=
b1l<-bootdist(f1l)
plot(b1l)
summary(b1l)
@

@@

<<descdistx2,fig=TRUE>>=
descdist(x2,discrete=T,boot=1000)
@

@@

Below is the result of the fit of a Poisson distribution with the bootstrap simulations.
<<fitdistx2p,fig=TRUE>>=
f2p<-fitdist(x2,"pois")
plot(f2p)
summary(f2p)
gofstat(f2p,print.test=TRUE)
b2p<-bootdist(f2p)
summary(b2p)
@

Below is the result of the fit of a negative binomial distribution with the boostrap simulations.
<<fitdistx2nb,fig=TRUE>>=
f2n<-fitdist(x2,"nbinom")
plot(f2n)
summary(f2n)
gofstat(f2n,print.test=TRUE)
b2n<-bootdist(f2n)
summary(b2n)
@

From goodness-of-fit graphs, Chi-squared statistics, AIC and BIC values,
it seems better to choose the fit of a negative binomial distribution
for this dataset even it has one more parameter than the Poisson one. This was not obvious while
looking at the skewness-kurtosis graph. This graph must be used cautiously especially for continuous 
distributions far from the normal distribution or for discrete distributions. It is only indicative.
